---
title: 'Lab 4: RMarkdown'
author: "Noa Avitan"
date: "2023-02-15"
output: html_document
---

### Load and Check Data

##### 1. Load tidyverse library. 
Set the chunk option message=FALSE to avoid the loading messages to show up in the document.

```{r, message=FALSE}
library(tidyverse)
```


##### 2. Load the dataset. 
Remember that the working directory of the knitting process is the same as the folder where your rmarkdown file is located As above, ensure that the informative messages are not displayed in the document.

```{r, message=FALSE}
incomeheight <- read_delim("income-height.csv.bz2")
```


##### 3. How many observations and variables do you have in this dataset? 
Answer in a complete sentence like “This dataset contains 4 rows and 2 columns” using the inline code chunks.

      This dataset contains `r nrow(incomeheight)` rows and `r ncol(incomeheight)` columns.


##### 4. List the variable names of the dataset, separated by comma, using a complete sentence.
Hint: check out function str_flatten_comma() in stringr library.

     The names of all the variables in the dataset are: `r str_flatten_comma(names(incomeheight), last = ", and ")`.
     
##### 5. How many missings are there in variables income, height, age and sex? 
Again, answer in a complete sentence, injecting the numbers into the text like There are 5 missings in variable “age”, 4 missings in ...

      There are `r sum(is.na(incomeheight$income))` missings in variable _income_, `r sum(is.na(incomeheight$height))` missings in variable _height_, `r sum(is.na(incomeheight$age))` missings in variable _age_, and `r sum(is.na(incomeheight$sex))` missings in variable _sex_.

### Do taller people earn more?

##### 1. Find the average height in data. Answer in a complete sentence, round the result as appropriate.
```{r, message=FALSE}
avg_height <- incomeheight %>% 
  summarise(avg_height = mean(height)) %>% 
  round(2)
```
      The average height in the data is `r avg_height` inches.


##### 2. Compute the average income for those who are shorter than the average height and those who are taller than the average height. Do this in a code chunk and show the code and the output. Comment your finding in text below.
Hint: you can use group_by(), grouping by a logical condition like group_by(height >
avgHeight).
```{r}
below_income <- incomeheight %>% 
  mutate(avg_height = mean(height)) %>% 
  filter(height < avg_height) %>% 
  summarise(below_income = mean(income))
below_income
above_income <- incomeheight %>% 
  mutate(avg_height = mean(height)) %>% 
  filter(height > avg_height) %>% 
  summarise(above_income = mean(income))
above_income
```
      It seems like those who are above the average height tend to have a higher average income compared to those who are below the average height.

##### 3. But does the difference also hold if we do it separately by sex? Now compute average income separately for men and women who are shorter and taller than the corresponding male and female average.
Hint: you may first group by sex, then compute for every person if they are taller than the
average, and then group again, this time by sex and if they are taller. Finally use summarize()
to compute the average income.
```{r}
fem_bel_income <- incomeheight %>% 
  filter(sex == "female") %>% 
  mutate(avg_fem_height = mean(height)) %>% 
  filter(height < avg_fem_height) %>% 
  summarize(female_bel_income = mean(income))
fem_bel_income
fem_abv_income <- incomeheight %>% 
  filter(sex == "female") %>% 
  mutate(avg_fem_height = mean(height)) %>% 
  filter(height > avg_fem_height) %>% 
  summarize(female_abv_income = mean(income))
fem_abv_income
mal_bel_income <- incomeheight %>% 
  filter(sex == "male") %>% 
  mutate(avg_mal_height = mean(height)) %>% 
  filter(height < avg_mal_height) %>% 
  summarize(male_bel_income = mean(income))
mal_bel_income
mal_abv_income <- incomeheight %>% 
  filter(sex == "male") %>% 
  mutate(avg_mal_height = mean(height)) %>% 
  filter(height > avg_mal_height) %>% 
  summarize(male_abv_income = mean(income))
mal_abv_income
```

##### 4. Now comment your results in a sentence–something like “short women earn $27,000 while tall women earn $32,000 in average...”. However, the numbers should be extracted from your results above using inline code chunks, not just copied in the text.
Do not forget to round the numbers as appropriate! You may also check out the function prettyNum for better printing of large numbers.

      Women who are below average female height earn $`r prettyNum(fem_bel_income)` on average, while women who are above the average female height earn $`r prettyNum(fem_abv_income)` on average.
      Men who are below average male height earn $`r prettyNum(mal_bel_income)` on average, while men who are above the average male height earn $`r prettyNum(mal_abv_income)` on average.



##### 5. Make a scatterplot where you plot height versus income. Do not forget to add informative labels. Comment the plot in text.
Hint: if using base-R plotting then you can use pch="." to make the scatterplot of dots, not circles.
```{r}
plot(incomeheight$height, incomeheight$income, xlab = "Height (Inches)", ylab = "Income ($)", main = "Height versus Income")
```

